---
title: "p8105_hw2_aar2192"
author: "Amadeia Rector"
date: "9/28/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
#Problem 1

###**Reading** and **cleaning** the data

```{r Read_clean_data}
nyc_transit =
  read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, station_name:entry,vending, ada) %>% 
  mutate(entry = recode(entry,"YES" = TRUE, "NO" = FALSE))
```

### Describing the Data
```{r Describing_the_data}
str(nyc_transit)
head(nyc_transit)
```
The dataset contains **1868 observations** and **19 variables**: line, station_name, station_latitude, station_longitude, route1, route2, route3, route4, route5, route6, route7, route8, route9, route10, route11, entrance_type, entry, vending, ada. The variable types are **character**, **integer**, and **logical**.

The data were first cleaned removing spaces between words that were then instead turned into underscores to join words, and letters were decapitalized. Then, certain variables were retained and one variable, _entry_, was recategorized as logical. The **dimensions** of the dataset are **6X19**.

The data do **_not_** appear to be tidy as certain columns, for example route1-route11 are values rather than separate variables and thus should be regrouped in a tidier format.






