---
title: "p8105_hw2_aar2192"
author: "Amadeia Rector"
date: "9/28/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
#Problem 1

###**Reading** and **cleaning** the data

```{r Read_clean_data}
nyc_transit =
  read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, station_name:entry,vending, ada) %>% 
  mutate(entry = recode(entry,"YES" = TRUE, "NO" = FALSE))
```

### Describing the Data
```{r Describing_the_data}
nyc_transit
str(nyc_transit)
```
The dataset contains **1868 observations** and **19 variables**: line, station_name, station_latitude, station_longitude, route1, route2, route3, route4, route5, route6, route7, route8, route9, route10, route11, entrance_type, entry, vending, ada. The variable types are **character**, **integer**, and **logical**.

The data were first cleaned removing spaces between words that were then instead turned into underscores to join words, and letters were decapitalized. Then, certain variables were retained and one variable, _entry_, was recategorized as logical. As mentioned previously, the **dimensions** of the dataset are **1,868 X 19**.

The data do **_not_** appear to be tidy as certain columns, for example route1-route11 are values rather than separate variables and thus should be regrouped in a tidier format.

### Answering questions

The following are questions regarding the nature of this dataset.

#### How many distinct stations are there?
```{r distinct_stations}
nyc_transit %>% 
  distinct(station_name)
```

There are 356 unique station names.

#### How many stations are ADA compliant?
```{r ada_compliant_stations}
summary(nyc_transit$ada==TRUE)
```
There are 468 ADA compliant stations.

#### What proportion of station entrances/exits without vending allow entrance?



